---
title: "Demos on logistic regression"
author: "Wing-Ho Ko"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages

```{r}
library(tidyverse)
library(broom)
library(statmod)
library(ecotox)
library(ciTools)
```

## Example 1: mites occurrence dataset

### Loading data

We will work with the `mites` dataset, which record whether mites of genus _Galumna_ is found in 70 sites of different environmental conditions.

```{r}
mites <- read_csv("data/mites.csv") |> mutate(
  Substrate = as.factor(Substrate),
  Shrub = as.factor(Shrub),
  Topo = as.factor(Topo)
)
glimpse(mites)
```

Our response variable in this case is `pa`, and for the purpose of analysis we will look at how the presence/absence of _Galumna_ is affected by water content (`WatrCont`) and topography (`Topo`). 

```{r}
ggplot(mites) + geom_point(aes(x=WatrCont, y=pa, color=Topo, shape=Topo))
```

### Performing logistic regression

```{r}
mites_glm <- glm(pa ~ WatrCont * Topo, binomial, mites)
summary(mites_glm)
```

### Extracting information programatically

```{r}
glance(mites_glm)
```

```{r}
tidy(mites_glm)
```

```{r}
fitted(mites_glm)
```

```{r}
confint(mites_glm)
```

### Making predictions

Create new data frame for predictions

```{r}
water_new <- seq(130, 830, 20)
topo_new <- distinct(mites["Topo"])$Topo

mites_new <- data.frame(
  WatrCont = rep(water_new, each=length(topo_new)),
  Topo = rep(topo_new, times=length(water_new))
)

```

Making predictions

```{r}
mites_pred <- predict(mites_glm, mites_new, type="response")
mites_pred <- cbind(mites_new, mites_pred) |> rename (pa = mites_pred)
```

Plot the predictions against the data

```{r}
ggplot(mites, aes(x=WatrCont, y=pa, color=Topo)) + 
  geom_point(aes(shape=Topo)) + 
  geom_line(aes(linetype=Topo), mites_pred)
```

Use `add_ci()` to create both prediction and confidence intervals (on the response scale)

```{r}
mites_ci <- add_ci(mites_new, mites_glm)
mites_ci
```

Again plot the prediction against observation

```{r}
ggplot(mites, aes(x=WatrCont, y=pa, color=Topo)) + 
  geom_point(aes(shape=Topo)) + 
  geom_line(aes(y=pred, linetype=Topo), mites_ci) + 
  geom_ribbon(aes(y=NULL, ymin=LCB0.025, ymax=UCB0.975, fill=Topo), mites_ci, alpha=0.2)
```


## Example 2: mice survival dataset

An experiment studied the survival of mice after receiving a test dose of culture with five different doses of antipneumococcus serum (in cc). The data is loaded below:

```{r}
serum <- read_csv("data/serum.csv") |> mutate(
  Survivors = as.integer(Survivors),
  Number = as.integer(Number)
)
```

```{r}
glimpse(serum)
```

### Fitting a logistic model

Add a `Deceased` column to the data

```{r}
serum <- serum |> mutate(Deceased = Number - Survivors)
```

Perform fit and print summary

```{r}
serum_glm <- glm(cbind(Survivors, Deceased) ~ log10(Dose), binomial, serum)
summary(serum_glm)
```
Predict confidence interval

```{r}
serum_ci <- add_ci(data.frame(Dose=serum$Dose), serum_glm)
serum_ci
```

Plot the fitted value (with confidence interval) with the effective dose (with confidence interval) and the raw data

```{r}
ggplot() + geom_point(aes(x=Dose, y=Survivors/Number), serum) +
  geom_line(aes(x=Dose, y=fitted(serum_glm)), serum, linetype="dashed") + 
  geom_ribbon(
    aes(x=Dose, ymin=LCB0.025, ymax=UCB0.975), 
    serum_ci, color="gray", alpha=0.2
  ) + 
  scale_x_continuous(transform="log10")
```


## Example 3: swiss banknotes redux

```{r}
banknotes <- read_csv("data/banknotes.csv") |>
  mutate(Currency = as.integer(Currency))
```

```{r}
bnotes_glm <- glm(
  Currency ~ Length + Left + Right + Bottom + Top, 
  binomial, banknotes
)

summary(bnotes_glm)
```

```{r}
ggplot(banknotes, aes(x=Bottom, y=Top, color=as.factor(Currency))) + 
  geom_point()
```

```{r}
bnotes_glm <- glm(
  Currency ~ Bottom + Top, 
  binomial, banknotes
)

summary(bnotes_glm)
```

```{r}
banknotes |> summarize(
  min_Bottom = min(Bottom),
  max_Bottom = max(Bottom),
  min_Top = min(Top),
  max_Top = max(Top)
)
```

```{r}
new_Bottom <- seq(7, 13, 0.1)
new_Top <- seq(7, 13, 0.1)
new_bnotes <- data.frame(
  Bottom = rep(new_Bottom, each=length(new_Top)),
  Top = rep(new_Top, times=length(new_Bottom))
)

bnotes_pred <- add_ci(new_bnotes, bnotes_glm)
```

```{r}
ggplot() + 
  geom_contour(
    aes(x=Bottom, y=Top, z=pred), 
    bnotes_pred, 
    breaks = c(0.1, 0.3, 0.5, 0.7, 0.9),
    linetype="dashed"
  ) + 
  geom_point(
    aes(x=Bottom, y=Top, color=as.factor(Currency)),
    banknotes
  )  
```

```{r}
bnotes_fitted <- banknotes |> 
  mutate(class = ifelse(fitted(bnotes_glm) > 0.5, 1, 0))
```

```{r}
bnotes_fitted |> summarize(
  correct = sum(class == Currency),
  incorrect = sum(class != Currency),
  accuracy = mean(class == Currency)
)
```


## Example 4: model selection and diagonoses

Work again with the mites dataset:

```{r}
mites <- read_csv("data/mites.csv") |> mutate(
  Substrate = as.factor(Substrate),
  Shrub = as.factor(Shrub),
  Topo = as.factor(Topo)
)
glimpse(mites)
```

### Model selection

Start with the model with full interactions:

```{r}
mites_glm1 <- glm(pa ~ WatrCont * Topo, binomial, mites)
```

Compare with the model for which the effects of the explanatory variables are additive:

```{r}
mites_glm2 <- glm(pa ~ WatrCont + Topo, binomial, mites)
anova(mites_glm2, mites_glm1)
```

Thus we should prefer the simpler model.

Alternatively, the `step()` function can be used for stepwise model selection:

```{r}
mites_glm_best <- step(mites_glm1, direction="backward")
```

We see that the stepwise selection picked the same model with additive terms

### Model diagnosis

Construct the quantile residuals

```{r}
mites_qresid <- qresid(mites_glm_best)
mites_fitted <- fitted(mites_glm_best)
```

Plot the quantile residuals against the fitted values

```{r}
ggplot() + geom_point(aes(x=mites_fitted, y=mites_qresid))
```

The plot looks OK (no systematic trend in mean or variance). Now look at the Q-Q plot

```{r}
ggplot(mapping=aes(sample=mites_qresid)) + geom_qq() + geom_qq_line()
```

Finally check the variance

```{r}
var(mites_qresid)
```

In conclusion, we do not see any serious problems

